Effect of quality-adjusted production function inputs

### 2015-04-14 Articles/Presentation_ESEE_2015/ESEE 2015.Rmd
Changes to ESEE playground.
Restructured the ESEE work around questions.

Fitting is accomplished with \emph{formulas} of the form

@@

Note that 
`scale`   $= \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the `formula` is specified as \texttt{y~$\sim$~k~+~l~+~e~+~time}.

@@

Note that `scale` $= \gamma$. 
Equations for $\alpha_1$, $\alpha_2$, and $\alpha_3$ are given in Table \ref{tab:CES_abg_equations}.

@@

# Cobb-Douglas results

```{r CD unadjusted}
naturalCoef(cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                    data = dataPT, constrained = TRUE))
naturalCoef(cdModel(formula = iGDP ~ iKstkO.WwithRD + iL + iXp + iYear, 
                    data = dataUK, constrained = TRUE))
```

```{r CD quality adjusted}
naturalCoef(cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                    data = dataPT, constrained = TRUE))
naturalCoef(cdModel(formula = iGDP ~ iKservO.WwithRD + ihLest + iU + iYear, 
                    data = dataUK, constrained = TRUE))
```

>>>>

Fitting is accomplished with `formula`s of the form

@@

Note that 
`scale`  $= \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the `formula` is specified as \texttt{y~$\sim$~k~+~l~+~e~+~time}.
$\theta$ is expected to be near unity.

@@

Note that `scale` $= \gamma$ and is expected to be close to unity. 
Equations for $\alpha_1$, $\alpha_2$, and $\alpha_3$ are given in Table \ref{tab:CES_abg_equations}.

@@

# Questions

## What effect does including energy and quality-adjustment have on the Solow residual in the Cobb-Douglas model?

First, use non-quality-adjusted data for Portugal and the Cobb-Douglas model.

```{r reduce Solow residual PT unadjusted}
ptCDunadjNoE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
                        data = dataPT, constrained = TRUE)
ptCDunadjWithE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
naturalCoef(ptCDunadjWithE)
```

The Solow residual (`lambda`) decreased from 
`r naturalCoef(ptCDunadjNoE)$lambda` to 
`r naturalCoef(ptCDunadjWithE)$lambda`,
so it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the Solow residual significantly.

What about for quality-adjusted data and Portugal?

``` {r reduce Solow residual PT adjusted}
ptCDadjNoE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
                        data = dataPT, constrained = TRUE)
ptCDadjWithE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
naturalCoef(ptCDadjWithE)
```

Using quality-adjusted inputs, the Solow residual (`lambda`) moves from 
`r naturalCoef(ptCDadjNoE)$lambda` to
`r naturalCoef(ptCDadjWithE)$lambda`.

When we look at the effect of both quality-corrected factors of production and energy,
we see that both including energy and quality-adjusting the factors of production 
reduce the Solow residual (`lambda`). 
Quality-adjustment gets part of the way toward eliminating the Solow residual.
Including energy (whether or not the factors of production are quality-adjusted)
reduces the Solow residual significantly.

Energy/Adjusted | Unadjusted                             | Quality-adjusted
--------------- | -------------------------------------- | -----------------------------------
Without energy  | `r naturalCoef(ptCDunadjNoE)$lambda`   | `r naturalCoef(ptCDadjNoE)$lambda`
With energy     | `r naturalCoef(ptCDunadjWithE)$lambda` | `r naturalCoef(ptCDadjWithE)$lambda`

Note that the contribution of capital (`alpha_1`)
always decreases as energy (`alpha_3`) is allowed to play a role.

What about for the UK?

```{r reduce Solow residual UK}
ukCDunadjNoE <- cdModel(formula = iGDP ~ iKstkO.WwithRD + iL + iYear, 
                        data = dataUK, constrained = TRUE)
ukCDunadjWithE <- cdModel(formula = iGDP ~ iKstkO.WwithRD + iL + iXp + iYear, 
                          data = dataUK, constrained = TRUE)
ukCDadjNoE <- cdModel(formula = iGDP ~ iKservO.WwithRD + ihLest + iYear, 
                        data = dataUK, constrained = TRUE)
ukCDadjWithE <- cdModel(formula = iGDP ~ iKservO.WwithRD + ihLest + iU + iYear, 
                          data = dataUK, constrained = TRUE)
```

Energy/Adjusted | Unadjusted                             | Quality-adjusted
--------------- | -------------------------------------- | -----------------------------------
Without energy  | `r naturalCoef(ukCDunadjNoE)$lambda`   | `r naturalCoef(ukCDadjNoE)$lambda`
With energy     | `r naturalCoef(ukCDunadjWithE)$lambda` | `r naturalCoef(ukCDadjWithE)$lambda`

For the UK, the quality-adjusted, with energy Solow residual is 
`r naturalCoef(ukCDadjWithE)$lambda * 100`%,
a very small number indeed.
For the UK, it appears that both quality-adjustment and addition of energy as a factor of production
can effectively eliminate the Solow residual,
reducing it to `r naturalCoef(ukCDadjWithE)$lambda`.

Interestingly, the Solow residual acutally increases from 
`r naturalCoef(ukCDunadjNoE)$lambda` to
`r naturalCoef(ukCDunadjWithE)$lambda`
for unadjusted factors of production when energy is added.
To find out why, let's look at all the coefficients.

``` {r}
naturalCoef(ukCDunadjNoE)
naturalCoef(ukCDunadjWithE)
```

When energy is included, much of the effect of capital (`alpha_1`) is 
attributed to energy (`alpha_3`), 
as `alpha_1` moves from 
`r naturalCoef(ukCDunadjNoE)$alpha_1` to
`r naturalCoef(ukCDunadjWithE)$alpha_1`.

When we look at quality-adjusted inputs, we find similar results.

``` {r}
naturalCoef(ukCDadjNoE)
naturalCoef(ukCDadjWithE)
```

`alpha_1` moves from 
`r naturalCoef(ukCDadjNoE)$alpha_1` to
`r naturalCoef(ukCDadjWithE)$alpha_1`.
